---
title: "Distributional Consequences of Climate Policy"
subtitle: "ESM 204 - Assignment 3"
author: "Alex Gaynor & Carlos Simms"
output:
  html_document:
    code_folding: hide
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE,
                      results = FALSE)

# Setup:

rm(list=ls())

options(scipen=999) # Gives normal numbers, not scientific ones.

gc() # Cleans ram.

# Packages:

packages=c("dplyr",
           "tidyr",
           "stringr",  
           "ggplot2",
           "stargazer",
           "cowplot",
           "janitor") 

# Fancy way to load packages (if its installed, run it, otherwise install it):

lapply(1:length(packages), 
       function(x)
         ifelse((require(packages[x],
                         character.only=TRUE)==FALSE),
                install.packages(packages[x]),
                require(packages[x],
                        character.only=TRUE)))
         
# Set directory (sets the directory where the script is saved to avoid complicated directory paths):

setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) 

getwd()
```

```{r}
# Load data (useful if letters are in the data to not let R do funny stuff):

gas_demand_data <- read.csv("HW3_data.csv", 
                            stringsAsFactors = F)

gas_demand <- gas_demand_data %>% 
  clean_names %>% 
  rename("price" ="price_dollars", 
         "q_low" = "q_low_gallons", 
         "q_high" = "q_high_gallons")

# Plot data:

plot_1 <- ggplot(data=gas_demand, 
                 aes(y=price)) +
  geom_point(aes(x=q_low)) +
  geom_point(aes(x=q_high)) +
  theme_cowplot(16)

# Estimate linear model (model has an implied intercept):

model_demand_low <- lm(price ~ q_low, 
                       data=gas_demand)

a_low <- model_demand_low$coefficients[[1]]

b_low <- model_demand_low$coefficients[[2]]

model_demand_high <- lm(price ~ q_high, 
                        data=gas_demand)

a_high <- model_demand_high$coefficients[[1]]

b_high <- model_demand_high$coefficients[[2]]

price_fitted_low <- a_low + b_low*gas_demand$q_low

price_fitted_high <- a_high + b_high*gas_demand$q_high

plot_2 <- plot_1 +
  geom_line(aes(y=price_fitted_low, 
                x=q_low)) +
  geom_line(aes(y=price_fitted_high, 
                x=q_high))

# Low income demand function: `P = `r b_low`Q + `r a_low``

# High income demand function: `P = `r b_high`Q + `r a_high``

# Inverse demand:

inverse_demand <- function(q, 
                           model){
  p <- model$coefficients[[1]] + model$coefficients[[2]]*q
  return(p)
  }

demand <- function(p, 
                   model){
  q <- (p - model$coefficients[[1]])/model$coefficients[[2]]
  return(q)
  }

demand(0,
       model_demand_low)

inverse_demand(176746.5,
               model_demand_low)

inverse_demand(demand(0,
                      model_demand_low),
               model_demand_low)

demand(3,
       model_demand_high)

inverse_demand(176746.5,
               model_demand_low)

inverse_demand(demand(0,
                      model_demand_low),
               model_demand_low)

a_low_invert <- demand(0,
                       model_demand_low)

a_high_invert <- demand(0,
                        model_demand_high)

# Determine the inverse slope:

b_low_invert <- 1/model_demand_low$coefficients[[2]]

b_high_invert <- 1/model_demand_high$coefficients[[2]]

# Determine the aggregate:

a_tot_invert <- a_low_invert + a_high_invert

b_tot_invert <- b_low_invert + b_high_invert

a_tot <- -a_tot_invert/b_tot_invert

b_tot <- 1/b_tot_invert

demand_alt <- function(p){
  q <- (p - a_tot)/b_tot
  return(q)
  }

demand_alt(0)

# Determine the marginal cost curve:

s_slope <- 3/demand_alt(3)

# $51 per metric ton/2204.62 lbs per metric ton * 19.6 lbs/gallon gas.

t = (51/2204.62)*19.6
```

# 1) Marginal Externality Cost of Gasoline

The marginal externality cost per gallon of gasoline is `$``r round(t, 2)` and found by `$51 per metric ton/2204.62 lbs per metric ton * 19.6 lbs per gallon gas`.

```{r}
# Social benefit:

social_low = (a_low-3)*(demand(3,model_demand_low))/2

social_high = (a_high-3)*(demand(3,model_demand_high))/2

social_tot <- social_low + social_high

high_percentage <- 100*social_high/social_tot

# Producer benefit:

producer <- (3)*(demand_alt(3))/2

# Cost for environment:

env_cost <- t*(demand_alt(3))

# Plotting the aggregate demand and supply:

plot_e <- plot_2 + xlim(0, 
                        800000) +
  ylim(0,16) +
  geom_abline(intercept=a_tot, 
              slope=b_tot, 
              color = "green") +
  geom_abline(intercept=0, 
              slope=s_slope, 
              color = "blue")
```

# 2) Curves

### What is the aggregate daily demand curve for gasoline?
From $0 to `r round(a_low,2)`: P = `r round(b_tot,7)`Q + `r round(a_tot,2)`, from `r round(a_low,2)` to infity:  P = `r round(b_high,7)`Q + `r round(a_high,2)`  

### What is the supply curve for gasoline?
P = `r round(s_slope,8)`Q + 0

### What is the “benefit” to consumers under the status quo?
`r round(social_tot,1)`

### What is the “benefit” to producers under the status quo?
`r round(producer,1)`

### What is the environmental cost under the status quo?
`r round(env_cost,1)`

# 3) Consumer Benefit

### How is the current consumer benefit divided between “high” and “low” income consumers?
"High" income consumers attain `$``r round(social_high,0)` of benefit and "low" income consumers  attain `$``r round(social_low,0)` of benefit for a total of `$``r round(social_tot,0)`. "High" income consumers receive `r round(high_percentage,0)` percent of the total benefit.

# 4) Derive the Optimal Gas Tax: `$``r round(t,2)` 

### a) How will this tax affect the amount of gasoline produced and consumed?
The amount of gasoline produced and consumed will move from `r round(demand_alt(3), 0)`, down to `r round (demand_alt(3.45), 0)`. 

### b) How will this tax affect the price of gasoline?
The price of gas for consumers is $3.45 inclusive of the tax.
Meanwhile, suppliers get paid $3 per gallon.

```{r}
# Social benefit:

t_var <- function(x){
  y <- (x/2204.62)*19.6
  return(y)
  }

t_var(51)

wellfare_low <- function(x){
  y <- (x/2204.62)*19.6
  z <- ((a_low-(3+y))*(demand((3+y),model_demand_low))/2)
  return(z)
  }

wellfare_low(51)

wellfare_high <- function(x){
  y <- (x/2204.62)*19.6
  z <- ((a_high-(3+y))*(demand((3+y),model_demand_high))/2)
  return(z)
  }

wellfare_high(51)

# New producer surplus:

producer_surplus <-function(x){
  y <- (x/2204.62)*19.6
  z <- (s_slope*(demand_alt(3+y))*(demand_alt(3+y))/2)
  return(z)
  }

producer_surplus(51)

# Tax revenue and environemntal damage:

tax_revenue <-function(x){
  y <- (x/2204.62)*19.6
  z <- (demand_alt(3+y)*y)
  return(z)
  }

tax_revenue(51)

net_high <- social_high - wellfare_high(51)

total_welfare_low <- social_low - tax_revenue(51)

net_low <- total_welfare_low - wellfare_low(51)

net_producer <- producer - producer_surplus(51)
```

### c) How will this tax affect the overall welfare of “high” income consumers?
Welfare for "high" income consumers will go from `r round(social_high,0)` down to `r round(wellfare_high(51),0)`, or a loss of `r round(net_high,0)`.

### d) How will this tax affect the overall welfare of “low” income consumers?
Before the tax, total welfare for "low" income consumers `r round(total_welfare_low,0)` was the total benefit `r round(social_low,0)` for "low" income people - the environmental damage burden that fell on them `r round(tax_revenue(51),0)`. Considering that no environemtnal damage occurs after the tax, the new total benefit for "low" income people is `r round(wellfare_low(51),0)`, or a gain of `r round(-net_low,0)`.

### e) How will this tax affect gas producers?
Supplier surplus for producers will go from `r round(producer,0)` down to `r round(producer_surplus(51),0)`, or a loss of `r round(net_producer,0)`.

### f) How will this tax affect total environmental damage?
At this tax rate there is no environmental damage accounted for with a social cost of carbon of $51.

### g) How will this tax affect total tax revenue generated?
Instead of $0 in tax, this will generate `r round(tax_revenue(51),0)`.

```{r}
# Redistributed:

pretax_perc_high <- demand(3,
                           model_demand_high)/demand_alt(3)

# y = tax rate determined by scc input, z = h consumer benefit at $3 + tax rate, j = (tax revenue of # gallons consumed at optimal consumption with new tax rate) multiplied by percentage of consumption by h consumers before tax, w = 

redist_h <- function (x){
  y <- (x/2204.62)*19.6
  z <- ((a_high-(3+y))*(demand((3+y),
                               model_demand_high))/2)
  j <- demand_alt(3+y)*y*pretax_perc_high
  w <- z+j
  return(w)
  }

redist_l <- function (x){
  y <- (x/2204.62)*19.6
  z <- ((a_low-(3+y))*(demand((3+y),
                              model_demand_low))/2)
  j <- demand_alt(3+y)*y*(1/pretax_perc_high)
  w <- z+j
  return(w)
  }

benefit_tible <- tibble(x = c(51, 75, 100, 125, 150),
  y <- (x/2204.62)*19.6,
  q <- ((a_high-(3+y))*(demand((3+y),
                               model_demand_high))/2),
  z <- ((a_low-(3+y))*(demand((3+y),
                              model_demand_low))/2),
  v <- q + demand_alt(3+y)*y*pretax_perc_high,
  w <- z+demand_alt(3+y)*y*(1/pretax_perc_high),
  p <- producer_surplus(x))
```




